\bf Summary

### 2025-02-13 paper/paper.qmd
Add hard code author names into `.qmd` for in-text named citations so it renders properly in `.tex`

`traveltime` provides a spatial interface using object classes from the `terra` package [@terra]. It accepts points as `matrix`, `data.frame`, or `SpatVector` class objects; and the area of interest as an extent in the form of a `numeric`, `SpatExtent`, `SpatVector`, or `SpatRaster` class object; and returns the result as a `SpatRaster` class object. The travel time is calculated as movement over a resistance 'friction surface' [@gdistance2017]. To provide easy access to the existing friction surfaces generated by @weiss2020global, `traveltime` uses the package `malariaAtlas` to allow users to access friction surfaces for the area of interest; though users can also supply any other friction surfaces to `traveltime`. Functionality from `gdistance` [@gdistance2017] is also used internally to calculate the minimum least-cost-distance for each cell from the points of interest.

@@

Global maps of travel time to cities [@weiss2018global; @nelson2019suite] and health care facilities [@hulland2019travel; @weiss2020global] have generated significant interest and use[^1], and the city data set of @nelson2019suite is available to R users through the widely-used `geodata` package [@geodata]. There is clear demand for these type of products.

@@

@weiss2020global made their code available as an R script to allow for reproduction and extension of their analyses (<https://malariaatlas.org/wp-content/uploads/2022/11/R_generic_accessibilty_mapping_script_2020-1.txt> ). To further enable extension of this work, here we have developed an R package based on that code to seamlessly calculate the travel time from any arbitrary set of locations.

@@

```{r}
#| label: get-basemap
library(terra)
library(geodata)

@@

The `stations` data set included in the `traveltime` package is a 563 row, 2 column `matrix` containing the longitude (`x`) and latitude (`y`) of all LRT and MRT station exits in Singapore from @singdata:

@@

Now that we have the two items of data that we require, the next step is to prepare a friction surface for our area of interest. We will use the friction surface from @weiss2020global that can be downloaded by `traveltime` with the function `get_friction_surface()`. We can pass in our basemap `singapore_shapefile`, a `SpatVector`, directly as the `extent`. We're interested in walking time from a station, so we'll download the walking friction surface by specifying `surface = "walk2020"`. (Alternatively, we could use `surface = "motor2020"` for motorised travel if that were of interest.) We're also only interested in walking *on land* so we then mask out areas outside of the land boundary in `singapore_shapefile`:

>>>>

`traveltime` provides a spatial interface using object classes from the `terra` package [@terra]. It accepts points as `matrix`, `data.frame`, or `SpatVector` class objects; and the area of interest as an extent in the form of a `numeric`, `SpatExtent`, `SpatVector`, or `SpatRaster` class object; and returns the result as a `SpatRaster` class object. The travel time is calculated as movement over a resistance 'friction surface' [@gdistance2017]. To provide easy access to the existing friction surfaces generated by Weiss et al. [@weiss2020global], `traveltime` uses the package `malariaAtlas` to allow users to access friction surfaces for the area of interest; though users can also supply any other friction surfaces to `traveltime`. Functionality from `gdistance` [@gdistance2017] is also used internally to calculate the minimum least-cost-distance for each cell from the points of interest.

@@

Global maps of travel time to cities [@weiss2018global; @nelson2019suite] and health care facilities [@hulland2019travel; @weiss2020global] have generated significant interest and use[^1], and the city data set of Nelson et al. [@nelson2019suite] is available to R users through the widely-used `geodata` package [@geodata]. There is clear demand for these type of products.

@@

Weiss et al. [@weiss2020global] made their code available as an R script to allow for reproduction and extension of their analyses (<https://malariaatlas.org/wp-content/uploads/2022/11/R_generic_accessibilty_mapping_script_2020-1.txt> ). To further enable extension of this work, here we have developed an R package based on that code to seamlessly calculate the travel time from any arbitrary set of locations.

@@

```{r}
#| echo: true
#| label: get-basemap
library(terra)
library(geodata)

@@

The `stations` data set included in the `traveltime` package is a 563 row, 2 column `matrix` containing the longitude (`x`) and latitude (`y`) of all LRT and MRT station exits in Singapore from Land Transport Authority @singdata:

@@

Now that we have the two items of data that we require, the next step is to prepare a friction surface for our area of interest. We will use the friction surface from Weiss et al. [@weiss2020global] that can be downloaded by `traveltime` with the function `get_friction_surface()`. We can pass in our basemap `singapore_shapefile`, a `SpatVector`, directly as the `extent`. We're interested in walking time from a station, so we'll download the walking friction surface by specifying `surface = "walk2020"`. (Alternatively, we could use `surface = "motor2020"` for motorised travel if that were of interest.) We're also only interested in walking *on land* so we then mask out areas outside of the land boundary in `singapore_shapefile`:

### 2025-02-13 paper/paper.tex
Add hard code author names into `.qmd` for in-text named citations so it renders properly in `.tex`

@@

The \texttt{stations} data set included in the \texttt{traveltime}
package is a 563 row, 2 column \texttt{matrix} containing the longitude
(\texttt{x}) and latitude (\texttt{y}) of all LRT and MRT station exits
in Singapore from Land Transport Authority \citep{singdata}:

>>>>

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(terra)}
\FunctionTok{library}\NormalTok{(geodata)}

\NormalTok{singapore\_shapefile }\OtherTok{\textless{}{-}} \FunctionTok{gadm}\NormalTok{(}
  \AttributeTok{country =} \StringTok{"Singapore"}\NormalTok{,}
  \AttributeTok{level =} \DecValTok{0}\NormalTok{,}
  \AttributeTok{path =} \FunctionTok{tempdir}\NormalTok{(),}
  \AttributeTok{resolution =} \DecValTok{2}
\NormalTok{)}

\NormalTok{singapore\_shapefile}
\end{Highlighting}
\end{Shaded}

@@

The \texttt{stations} data set included in the \texttt{traveltime}
package is a 563 row, 2 column \texttt{matrix} containing the longitude
(\texttt{x}) and latitude (\texttt{y}) of all LRT and MRT station exits
in Singapore from Land Transport Authority \citet{singdata}:
